---
title: "Biostat 234: Lab 3"
author: "Minsoo Kim"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(R2jags)
library(lattice)
set.seed(1234)

load("AddBurnin.RData")

# Function that summarizes posterior distribution
mysummary = function(invector) {
  c(mean = mean(invector), 
    sd = sd(invector), 
    quantile(invector, .025), 
    quantile(invector, .975),
    `P > 0` = sum(invector > 0) / length(invector))
} 

data <- read.table("data.txt")
colnames(data) <- c("death", "n",	"intercept",	"iss",	"rts", 
                    "age", "ti", "age * ti")

# Estimate prior for beta from six samples
Xp <- as.matrix(data[1:6, 3:8])
Yp <- data[1:6, 1]
np <- data[1:6, 2]

invXp <- solve(Xp)

Xobs <- as.matrix(data[7:306, 3:8])
Yobs <- data[7:306, 1]
nobs <- data[7:306, 2]

#sink("priors.txt")
#cat("
#  model{
#  	betas <- invXp %*% logitp[]
#  
#  	for (j in 1:6) {
#  		logitp[j] <- logit(pie[j])
#  	}
#  	pie[1] ~ dbeta(1.1, 8.5)
#  	pie[2] ~ dbeta(3.0, 11.0)
#  	pie[3] ~ dbeta(5.9, 1.7)
#  	pie[4] ~ dbeta(1.3, 12.9)
#  	pie[5] ~ dbeta(1.1, 4.9)
#  	pie[6] ~ dbeta(1.5, 5.5)
#  }
#  ", fill = TRUE)
#sink()

ex1.data <- list(invXp = invXp)
ex1.inits <- rep(list(list(pie = c(0.5, 0.5, 0.5, 0.5, 0.5, 0.5))), 5)
ex1.parameters <- c("betas", "pie[1:6]")

ex1.out <- jags(ex1.data, ex1.inits, ex1.parameters, "priors.txt", 
                n.chains = 5, n.iter = 11000, n.burnin = 0, n.thin = 2, DIC = F)

Output1 <- AddBurnin(ex1.out$BUGSoutput$sims.array, burnin = 1000, n.thin = 2)
#print(Output1$Burnin.Summary)

#sink("posteriors.txt")
#cat("
#  model{
#  	betas <- invXp %*% logitp[]
#
#  	for (j in 1:6) {
#  		logitp[j] <- logit(pie[j])
#  	}
#  	pie[1] ~ dbeta(1.1, 8.5)
#  	pie[2] ~ dbeta(3.0, 11.0)
#  	pie[3] ~ dbeta(5.9, 1.7)
#  	pie[4] ~ dbeta(1.3, 12.9)
#  	pie[5] ~ dbeta(1.1, 4.9)
#  	pie[6] ~ dbeta(1.5, 5.5)
#  	
#  	for(i in 1:T) {
#  	  y[i] ~ dbern(p[i])
#  	  p[i] <- ilogit(inprod(x[i, ], betas[]))
#  	}
#  }
#  ", fill = TRUE)
#sink()

ex2.data <- list(x = Xobs, y = Yobs, T = 300, invXp = invXp)
ex2.inits <- rep(list(list(pie = c(0.5, 0.5, 0.5, 0.5, 0.5, 0.5))), 5)
ex2.parameters <- c("betas", "pie[1:6]")

ex2.out <- jags(ex2.data, ex2.inits, ex2.parameters, "posteriors.txt", 
                n.chains = 5, n.iter = 11000, n.burnin = 0, n.thin = 2, DIC = F)

#Treat the first 1000 iterations as a burn in	
Output2 = AddBurnin(ex2.out$BUGSoutput$sims.array, burnin = 1000, n.thin = 2)

ex3.data <- list(x = Xobs, y = Yobs, T = 100, invXp = invXp)
ex3.inits <- rep(list(list(pie = c(0.5, 0.5, 0.5, 0.5, 0.5, 0.5))), 5)
ex3.parameters <- c("betas", "pie[1:6]")

ex3.out <- jags(ex3.data, ex3.inits, ex3.parameters, "posteriors.txt", 
                n.chains = 5, n.iter = 11000, n.burnin = 0, n.thin = 2, DIC = F)

#Treat the first 1000 iterations as a burn in	
Output3 = AddBurnin(ex3.out$BUGSoutput$sims.array, burnin = 1000, n.thin = 2)
```

## Question 1
```{r, out.width="50%"}
temp <- Output2$Burnin.sims.matrix
beta1 <- temp[, 1]
beta2 <- temp[, 2]
beta3 <- temp[, 3]
beta4 <- temp[, 4]
beta5 <- temp[, 5]
beta6 <- temp[, 6]

plot(acf(beta1, lag.max = 200))
plot(acf(beta2, lag.max = 200))
plot(acf(beta3, lag.max = 200))
plot(acf(beta4, lag.max = 200))
plot(acf(beta5, lag.max = 200))
plot(acf(beta6, lag.max = 200))
```
Autocorrelations hit zero for the 6 regression coefficients at different lags as seen above. By about 200 lags, all  autocorrelations are zero. 

```{r, out.width="50%"}
pie1 <- temp[, 7]
pie2 <- temp[, 8]
pie3 <- temp[, 9]
pie4 <- temp[, 10]
pie5 <- temp[, 11]
pie6 <- temp[, 12]

plot(acf(pie1, lag.max = 200))
plot(acf(pie2, lag.max = 200))
plot(acf(pie3, lag.max = 200))
plot(acf(pie4, lag.max = 200))
plot(acf(pie5, lag.max = 200))
plot(acf(pie6, lag.max = 200))
```
The beta autocorrelations are better in general than the 6 pi’s as seen above. 

\newpage

## Question 2
```{r, warning=FALSE, message=FALSE}
library(knitr)
df <- t(round(apply(temp, 2, mysummary), 2))
kable(df)

temp2 <- Output1$Burnin.sims.matrix
df <- data.frame(`beta1 (T = 300)` = beta1, 
                 `beta2 (T = 300)` = beta2, 
                 `beta3 (T = 300)` = beta3, 
                 `beta4 (T = 300)` = beta4, 
                 `beta5 (T = 300)` = beta5, 
                 `beta6 (T = 300)` = beta6, 
                 `beta1 (prior)` = temp2[, 1],
                 `beta2 (prior)` = temp2[, 2],
                 `beta3 (prior)` = temp2[, 3],
                 `beta4 (prior)` = temp2[, 4],
                 `beta5 (prior)` = temp2[, 5],
                 `beta6 (prior)` = temp2[, 6], check.names = FALSE)

library(tidyverse)
df %>% select(`beta1 (T = 300)`, `beta1 (prior)`) %>% gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta2 (T = 300)`, `beta1 (prior)`) %>% gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta3 (T = 300)`, `beta3 (prior)`) %>% gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta4 (T = 300)`, `beta4 (prior)`) %>% gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta5 (T = 300)`, `beta5 (prior)`) %>% gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta6 (T = 300)`, `beta6 (prior)`) %>% gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 
```

As above, posterior distribution has much smaller variance compared to prior distribution for all regression coefficients.

## Question 3
```{r}
temp3 <- Output3$Burnin.sims.matrix
df <- data.frame(df, 
                 `beta1 (T = 100)` = temp3[, 1],
                 `beta2 (T = 100)` = temp3[, 2],
                 `beta3 (T = 100)` = temp3[, 3],
                 `beta4 (T = 100)` = temp3[, 4],
                 `beta5 (T = 100)` = temp3[, 5],
                 `beta6 (T = 100)` = temp3[, 6], check.names = FALSE)

kable(t(round(apply(df, 2, mysummary), 2)))

df %>% select(`beta1 (T = 300)`, `beta1 (prior)`, `beta1 (T = 100)`) %>% 
  gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta2 (T = 300)`, `beta1 (prior)`, `beta2 (T = 100)`) %>% 
  gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta3 (T = 300)`, `beta3 (prior)`, `beta3 (T = 100)`) %>% 
  gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta4 (T = 300)`, `beta4 (prior)`, `beta4 (T = 100)`) %>% 
  gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta5 (T = 300)`, `beta5 (prior)`, `beta5 (T = 100)`) %>% 
  gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 

df %>% select(`beta6 (T = 300)`, `beta6 (prior)`, `beta6 (T = 100)`) %>% 
  gather(key = "beta", value = "coefficient") %>% 
  ggplot(aes(coefficient, color = beta)) + 
  geom_density() 
```

As expected, posterior distribution incorporating more data gives sharper estimates of regression coefficients, but `T = 100` is already fairly good. 

## Question 4
$\pi$'s from `JAGS` output are posterior samples of the probability of death for the six samples, which were originally used for prior calculation of `beta`, after pooling information from the data. It is essentially refined estimate after incorporating the data. 

## Question 5
```{r, echo=TRUE}
#sink("posteriors2.txt")
#cat("
#  model{
#  	betas <- invXp %*% logitp[]
#
#  	for (j in 1:6) {
#  		logitp[j] <- logit(pie[j])
#  	}
#  	pie[1] ~ dbeta(1.1, 8.5)
#  	pie[2] ~ dbeta(3.0, 11.0)
#  	pie[3] ~ dbeta(5.9, 1.7)
#  	pie[4] ~ dbeta(1.3, 12.9)
#  	pie[5] ~ dbeta(1.1, 4.9)
#  	pie[6] ~ dbeta(1.5, 5.5)
#  	
#  	for(i in 1:T) {
#  	  y[i] ~ dbern(p[i])
#  	  p[i] <- ilogit(inprod(x[i, ], betas[]))
#  	}
#  	
#  	futurefit1 <- ilogit(betas[1] + betas[2] * 2 + betas[3] * 7.55 + betas[4] * 25)
#  	futurefit2 <- ilogit(betas[1] + betas[2] * 11 + betas[3] * 7.8408 + betas[4] * 42 +
#  	  betas[5] + betas[6] * 42)
#   futurefit3 <- ilogit(betas[1] + betas[2] * 16 + betas[3] * 7.8408 + betas[4] * 80 +
#  	  betas[5] + betas[6] * 80)
#  }
#  ", fill = TRUE)
#sink()

ex4.data <- list(x = Xobs, y = Yobs, T = 300, invXp = invXp)
ex4.inits <- rep(list(list(pie = c(0.5, 0.5, 0.5, 0.5, 0.5, 0.5))), 5)
ex4.parameters <- c("betas", "pie[1:6]", 
                    "futurefit1", "futurefit2", "futurefit3")

ex4.out <- jags(ex4.data, ex4.inits, ex4.parameters, "posteriors2.txt", 
                n.chains = 5, n.iter = 11000, n.burnin = 0, n.thin = 2, DIC = F)

Output4 = AddBurnin(ex4.out$BUGSoutput$sims.array, burnin = 1000, n.thin = 2)

kable(t(round(apply(Output4$Burnin.sims.matrix[, 7:9], 2, mysummary), 2)))
```